---
title: R Basics
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(arrow)

# We will learn what this code does in Milestone 2
animals <- read_parquet("data/intermediate_files.parquet")

```

## R Basics

In this milestone, you'll become familiar with the `mdrd` data set. 


## Recreation

### Part 1

The data we will use in our investigation is saved as an R object named `mdrd`. Use R to inspect the contents of `mdrd`:

- Display the contents of `mdrd` 
- Use `nrow()` to find the number of observations in `mdrd`
- Use `mean()` to find the average glomerular filtration rate (GFR) 
- Use `count()` from the dplyr package to count the number of observations associated with each patient
- Use `summary()` to generate a summary of the data set

*Hint: you may want to look up the help pages for each of these functions.*

Write your code in the following chunk:


```{r}
#| label: recreation
# DO NOT INSTALL TIDYVERSE PACKAGES. Remember that all tidyverse packages have already been installed for you, but you will need to *load* the ones you want to use.
```

## Extension

Explore `mdrd` in a new way. Consider using a new function, or something that you can learn from a help page. Alternatively, try to learn more about the other variables in `mdrd`. Keep in mind that we will study how to make plots for the next milestone.

Write your extension code in the following chunk:



```{r}
#| label: extension

```

