---
title: "Week 6 Milestone Template Report"
format:
  html
embed-resources: true
---

```{r}
#| label: setup
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  tidyverse,
  dtplyr,
  gt,
  arrow,
  rmarkdown,
  lubridate,
  quarto, 
  arrow,
  zoo
  )

# data you will need for this report
events_formatted <- read_parquet(
  '../data/intermediate_files/events_formatted.parquet')

# filter this data for LAME and TRIM in the event variable


```

```{=html}

<!-- The objective of this milestone is to start the creation of a
template you can use for reproducible reports going forward. Most of
this milestone will focus on customizing the look of the report that you
can use for future reports. We will also explore of the lameness data.We will use this template in future milestones.
A few notes:  
1) The logo contains a clickable link 
2) to use an image put it in the image folder (the one in the report is called 
  logo_no_background.png)
3)Any data (including farm name) that is reported in the text should use an inline text code chunk -->
```

Run this chuck below. It will open the example report in your browser.  You should delete this chunk in your final report

```{r}
#| label: open-html-r
#| eval: false

# 1. Define the file path
file_path <- "images/milestone_6.htm"

# 2. Open the browser
browseURL(file_path)
```


Objective

Explore recorded lameness/trim protocols

Use the protocols_letters1 variable to get the information about the
    lameness/trim protocols used.

```{r}

# put your graph code here

```

NA exploration

```{r}


# put your code to explore NA's here and name it na_lame_event so the code
# below will work

# we will learn better ways to show tables later but for now this works
rmarkdown::paged_table(na_lame_events)

```

Extension

For this week extension create a template for your future reports

-   customize the report to use your business' logo

-   use a theme/font colour scheme you prefer
