---
title: "Milestone 5 - Dates and Strings"
format: html
date: "`r Sys.Date()`"
embed-resources: true
---

```{r}
library(tidyverse)
library(quarto)
library(arrow)
library(waldo)

solution<-read_parquet('data/intermediate_files/events_formatted.parquet')

# read in files-----------------

list_files <- list.files(here::here("data/event_files")) # folder name where event files are located

events_original <- NULL

# i=1
for (i in seq_along(list_files)) {
  df <- read_csv(
    here::here(
      "data/event_files",
      list_files[i]
    ),
    # reads in all data as character string
    col_types = cols(.default = "c")
  )
  
   events_original <- bind_rows(events_original, df)
}

```

## Dates

The original dairy comp data is read into your environment and named "events_original".
For consistency all columns are read in as characters.

Use your new skills formating dates to create a new column named date_event that is a real date from the original column named Date.

Then create new variables that provide floor dates by month, season, and year. Name these floordate_month, floordate_year, floordate_season.

```{r initial denominator generation, message==FALSE}
evnets_new <- events_original |> 
  mutate(date_event       = mdy(Date),
         floordate_month  = floor_date(date_event, "month"),
         floordate_year   = floor_date(date_event, "year"), 
         floordate_season = floor_date(date_event, "season"))

```

Do your new dates match the dates in the solution?


# Strings

Find all the columns in the solution that contain the word "remark".  

These are all based on the original Remark column in dairy comp. 

Can you recreate them?  

Hint: don't spend too much time trying to get exactly the same values. 
Just spend time thinking about how you could use them to clean up remarks in a way that would help you answer a question.


```{r, echo = TRUE}
names(solution) |>
  str_subset(regex("remark", ignore_case = TRUE))

solution_new <- solution |>
  mutate(remark_numbers1_1 = str_extract(remark, "\\d+"),
         remark_letters1_1 = str_extract(remark, "\\d+")) |>
  select(remark_numbers1, remark_numbers1_1)

solution_new
```

# Extension - Your Turn

Start with events_original.  

Use your wrangling skils to fix date formats, and simplify free text to answer a question on this dairy.

You can choose any question you want. 

Here are some examples:

-  Which drug is most commonly used for mastitis treatment and has this changed over time?
-  What percent of breedings use sexed semen?  (this requires a large wrangle with an external key)
-  What specific lameness diagnosis are there and how frequently do they occur? 

```{r, echo = TRUE}
events_new_1 <- events_original |> 
  mutate(date_event       = mdy(Date),
         floordate_year   = floor_date(date_event, "year")) |>
filter(LACT >=1,
Event == "MAST") |>
   group_by(floordate_year, Protocols) |>
   summarize(n = n()) |> ungroup() 

events_new_1 |>
  ggplot(aes(floordate_year, fill = Protocols)) +
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = scales::percent) +
# geom_text(
#     stat = "count",
#     aes(group = floordate_year, label = scales::percent(after_stat(prop), accuracy = 0.1)),
#     position = position_fill(vjust = 0.5),
#     size = 3) +
ylab("Proportion") +
  xlab("Year")

solution |>
  filter(event == "BRED")
  
```
