---
title: R Basics
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(arrow)
library(skimr)

# We will learn what this code does in Milestone 2
animals <- read_parquet("../data/intermediate_files/animals.parquet")%>%
  filter(!(is.na(age_left)))

```

## R Basics

In this milestone, you'll become familiar with the `animals` data set.

## Recreation

### Part 1

The data we will use in our investigation is saved as an R object named `animals`. Use R to inspect the contents of `animals`:

-   Display the contents of `animals`
-   Use `nrow()` to find the number of observations in `animals`. Name this "number_of_rows". (hint: you should see "number_of_rows" in your environment)
-   Use `mean()` to find the average age and animal left (hint: use the variable named "age_left").  Name this "average_age_left".
-   Use `count()` from the dplyr package to count the number of observations associated with each location, breed and sex (hint: variable names should be "breed", "sex", "location_list". Name this 'counts'.


*Hint: you may want to look up the help pages for each of these functions.*

Write your code in the following chunk:

```{r}
#| label: recreation

animals

number_of_rows<-nrow(animals)

age_left<-mean(animals$age_left)

counts<-count(breed, sex)

```


##Check your answers

To check your answers run this code
```{r}

solutions<-read_csv('solutions_week_01.csv')

```

Now summarize the data: 
-   Use `summary()` to generate a summary of the data set
-   Use `skim()` to see a different type of summary of the data set (warning: this function may take some time)

Write your code in the following chunk:

```{r}
#| label: recreation of summary

summary(animals)

skim(animals)
```


## Extension

Explore `animals` in a new way. Consider using a new function, or something that you can learn from a help page. Alternatively, try to learn more about the other variables in `animals`. Keep in mind that we will study how to make plots for the next milestone.

Write your extension code in the following chunk:

```{r}
#| label: extension

```
